The limits of self commitment and private paternalism \thanks{}

### 2022-01-26 old.tex
checkpoint before edits of tables and figues

Combining the forcing and the choice arms ---a unique feature of our experiment--- allows us to estimate the counterfactuals to answer the question of how much money is left on the table by giving choice to clients. We calculate that about \hl{80}\% of clients in the fee-choice group chose contracts that induced \textit{higher} financial cost. On average they spend an extra \$\hl{238} MXN, close to \hl{11}\% of the average loan value. In particular \hl{92}\% of those that chose the status-quo contract would have been better in the fee-commitment contract, while only \hl{8}\% of those who chose the fee-commitment contract would have been better in the status-quo one. This suggests that forcing clients into commitment contacts (i.e. private paternalism) may be beneficial for the overwhelming majority of clients. After seeing our results Lender $P$ introduced frequent payment contracts in their loans portfolio.

@@

We document two main results in the promise arms. First, demand for the commitment-promise contract is three times larger than for the commitment fee contract, with \hl{31}\% take-up. This difference between both choice arms suggests that clients are indeed understanding and taking into account the contractual parameters in their choice, and that the fee commitment was probably too strong from their point of view. In spite of this much higher take-up, the \textit{Promise-choice arm} had no effect on outcomes on average, or even for those who selected into promising. Second, being forced in to the \textit{Promise} contract also had zero effects on average and throughout the distribution on financing cost and pawn recovery. %\footnote{There is a decrease in financial cost in the 50th and 75th quantiles in the forced promise-commitment arm however.}
Promises did not provide enough commitment so as to be reflected in behavior.

@@

\vspace{.2in}
\noindent \textbf{Randomization.} Randomization was done at the branch-day level.  Each day the system chose which types of contracts were on offer that day in the branch, and therefore which contract the client signed. Branch personnel did not know which treatment would happen on which day. They were told that there were 5 different ``types of contract-days'' and that the system chose randomly which applied on any given date, and that it could happen for instance that two consecutive dates had the same contract. They were also told that this way of operating was in place in various Lender $P$'s branches (they did not know which ones), and that it could be in place for several months. Randomizing at the day level limits the problem of contamination arising from clients realizing that other clients get different contracts than theirs. It also limits potential manipulation by appraisers, who in the presence of individual level randomization could potentially pick their preferred costumers from the line or tell them to wait until their desired contract shows up on the screen. Intra-branch day correlation on the probability of default (ICC) is small, at \hl{0.08}, so we don't lose much power vis-a-vis individual level randomization.

@@

How does overconfidence map to cost outcomes? Theory predicts that the more overconfident (naive) may be at the same time those that benefit more from commitment. Extrapolating individual treatment effects of the fee-forcing contract vs the status quo one to the choice arm (see below for more detail), and using them as a dependent variable we find that overconfident borrowers benefit \textit{more} from the commitment contract, with \hl{24.8}\% (=\$\hl{30}/\$\hl{121}) more savings (columns 5 and 6). Thus while we predict the overconfident benefit more from commitment, they demand less of it.

@@

Since the prediction of the treatment effect is an estimate and has noise, we do two things. First, we normalize $\widehat{TE}^{fee}_{i}$ by the value of the loan of person $i$ and introduce a threshold parameter $z$ and say that a choice is a ``mistake'' if the alternative contract would generate greater than $z$ in savings. That is, we introduce a threshold that we vary from 0\% to 15\% of the value of the loan. \hl{Second, we take into account uncertainty in estimation of treatment effects and report results using the extremes of the 95\% confidence interval bounds for $\widehat{TE}^{fee}_{i}$.}

Figure \ref{choose_wrong}(a) plots on the left Y-axis the fraction of clients that made a choice ``mistake'' and on the right Y-axis its size in pesos. On the X-axis it varies the threshold $z$. When $z=0$, we calculate that about 75\% of clients in the fee-choice group chose contracts that induced \textit{higher} financial cost.  %(90\% confidence interval [45,94]).
Most mistakes are failures to choose the fee-commitment contract (\hl{85}\%), while only \hl{15}\% of those who chose the fee-commitment contract would have been better in the status-quo one. The average mistake in pesos is \$250. %(90\% confidence interval [140,390]). 
Figure \ref{choose_wrong}(b) plots the same curve, but separately for overconfident and non-overconfident clients, using the definition of overconfidence in Section \ref{behavioral}. \hl{It shows that there are about 11 percentage points more mistakes in the overconfident population (when $z=0$), an important difference.}

@@

\vspace{.1in}
\noindent \textbf{Demand}. It turns out that the demand for the promise-commitment contract  is higher than for the fee-commitment contract by a factor of 3: \hl{31}\% of clients offered a choice between the status quo contract and the promise-commitment contract chose commitment. This means that the pecuniary fee indeed limited demand for commitment. In passing this means also that clients are responsive to contract terms and seem to understand the conditions. Interestingly, the set of people who demand the promise-commitment contract are different from those that demanded the fee-commitment one. While in both cases the economically vulnerable demand less, in the case of the promise-commitment those that we classify as present biased, those that report that they frequently fall into temptation buying, and those that ask for reminders of payment are \textit{less} likely to demand the promise-commitment contract.  This suggests that those taking up the promise commitment are not the ones needing the commitment (see Figure \ref{interactions_takeup} in the Appendix).

>>>>

Combining the forcing and the choice arms ---a unique feature of our experiment--- allows us to estimate the counterfactuals to answer the question of how much money is left on the table by giving choice to clients. We calculate that about 80\% of clients in the fee-choice group chose contracts that induced \textit{higher} financial cost. On average they spend an extra \$238 MXN, close to 11\% of the average loan value. In particular 92\% of those that chose the status-quo contract would have been better in the fee-commitment contract, while only 8\% of those who chose the fee-commitment contract would have been better in the status-quo one. This suggests that forcing clients into commitment contacts (i.e. private paternalism) may be beneficial for the overwhelming majority of clients. After seeing our results Lender $P$ introduced frequent payment contracts in their loans portfolio.

@@

We document two main results in the promise arms. First, demand for the commitment-promise contract is three times larger than for the commitment fee contract, with 31\% take-up. This difference between both choice arms suggests that clients are indeed understanding and taking into account the contractual parameters in their choice, and that the fee commitment was probably too strong from their point of view. In spite of this much higher take-up, the \textit{Promise-choice arm} had no effect on outcomes on average, or even for those who selected into promising. Second, being forced in to the \textit{Promise} contract also had zero effects on average and throughout the distribution on financing cost and pawn recovery. %\footnote{There is a decrease in financial cost in the 50th and 75th quantiles in the forced promise-commitment arm however.}
Promises did not provide enough commitment so as to be reflected in behavior.

@@

\vspace{.2in}
\noindent \textbf{Randomization.} Randomization was done at the branch-day level.  Each day the system chose which types of contracts were on offer that day in the branch, and therefore which contract the client signed. Branch personnel did not know which treatment would happen on which day. They were told that there were 5 different ``types of contract-days'' and that the system chose randomly which applied on any given date, and that it could happen for instance that two consecutive dates had the same contract. They were also told that this way of operating was in place in various Lender $P$'s branches (they did not know which ones), and that it could be in place for several months. Randomizing at the day level limits the problem of contamination arising from clients realizing that other clients get different contracts than theirs. It also limits potential manipulation by appraisers, who in the presence of individual level randomization could potentially pick their preferred costumers from the line or tell them to wait until their desired contract shows up on the screen. Intra-branch day correlation on the probability of default (ICC) is small, at 0.08, so we don't lose much power vis-a-vis individual level randomization.

@@

How does overconfidence map to cost outcomes? Theory predicts that the more overconfident (naive) may be at the same time those that benefit more from commitment. Extrapolating individual treatment effects of the fee-forcing contract vs the status quo one to the choice arm (see below for more detail), and using them as a dependent variable we find that overconfident borrowers benefit \textit{more} from the commitment contract, with 24.8\% (=\$30/\$121) more savings (columns 5 and 6). Thus while we predict the overconfident benefit more from commitment, they demand less of it.

@@

Since the prediction of the treatment effect is an estimate and has noise, we do two things. First, we normalize $\widehat{TE}^{fee}_{i}$ by the value of the loan of person $i$ and introduce a threshold parameter $z$ and say that a choice is a ``mistake'' if the alternative contract would generate greater than $z$ in savings. That is, we introduce a threshold that we vary from 0\% to 15\% of the value of the loan. Second, we take into account uncertainty in estimation of treatment effects and report results using the extremes of the 95\% confidence interval bounds for $\widehat{TE}^{fee}_{i}$.

Figure \ref{choose_wrong}(a) plots on the left Y-axis the fraction of clients that made a choice ``mistake'' and on the right Y-axis its size in pesos. On the X-axis it varies the threshold $z$. When $z=0$, we calculate that about 75\% of clients in the fee-choice group chose contracts that induced \textit{higher} financial cost.  %(90\% confidence interval [45,94]).
Most mistakes are failures to choose the fee-commitment contract (85\%), while only 15\% of those who chose the fee-commitment contract would have been better in the status-quo one. The average mistake in pesos is \$250. %(90\% confidence interval [140,390]). 
Figure \ref{choose_wrong}(b) plots the same curve, but separately for overconfident and non-overconfident clients, using the definition of overconfidence in Section \ref{behavioral}. It shows that there are about 11 percentage points more mistakes in the overconfident population (when $z=0$), an important difference.

@@

\vspace{.1in}
\noindent \textbf{Demand}. It turns out that the demand for the promise-commitment contract  is higher than for the fee-commitment contract by a factor of 3: 31\% of clients offered a choice between the status quo contract and the promise-commitment contract chose commitment. This means that the pecuniary fee indeed limited demand for commitment. In passing this means also that clients are responsive to contract terms and seem to understand the conditions. Interestingly, the set of people who demand the promise-commitment contract are different from those that demanded the fee-commitment one. While in both cases the economically vulnerable demand less, in the case of the promise-commitment those that we classify as present biased, those that report that they frequently fall into temptation buying, and those that ask for reminders of payment are \textit{less} likely to demand the promise-commitment contract.  This suggests that those taking up the promise commitment are not the ones needing the commitment (see Figure \ref{interactions_takeup} in the Appendix).
